---
output:
  pdf_document: default
  html_document: default
bibliography: references.bib
thanks: 'Code and data are available at: [https://github.com/Kluckyyou/STA313A2](https://github.com/Kluckyyou/STA313A2).'
---

<style>
/* CSS FOR INTRO SECTION */
.intro {
  height: 100vh;
  width: 100%;
  margin-bottom: 2em;
}

.intro-headers {
  position: absolute;
  top: 50%;
  left: 50%;
  transform: translate(-50%, -50%);
  text-align: center;
  color: lightskyblue;
  font-style: normal;
  width: 60%;
}

.title {
  font-size: 70px;                   /* Larger font size for the main title */
  font-weight: bold;                 /* Bold font weight for emphasis */
  color: lightskyblue;                    /* Yellow color similar to the example */
  font-family: Georgia, serif;       /* Use Georgia or a similar serif font */
  margin: 0;
  line-height: 1.1;                  /* Tighter line height */
  padding: 10px;                     /* Add padding around text */
  display: inline-block;             /* Keeps the border tight around text */
  text-shadow: 3px 3px 8px rgba(0, 0, 0, 0.7); /* Text shadow for emphasis */
  text-align: center;
}

.subtitle {
  font-size: 40px;            /* Slightly larger font size for the subtitle */
  font-weight: normal;        /* Normal font weight */
  color: lightskyblue;               /* White color for the subtitle */
  font-family: 'Times New Roman', Times, serif; /* A serif font for contrast */
  margin-top: 10px;
  line-height: 1.2;           /* Adjust line height for spacing */
  text-shadow: 3px 3px 8px rgba(0, 0, 0, 0.7); /* Text shadow for emphasis */
  text-align: center;
}

.author-date {
  font-size: 30px;
  font-weight: normal;
  color: lightblue;
  text-align: left;
  margin-top: 30px;
  font-style: italic;
  text-shadow: 3px 3px 8px rgba(0, 0, 0, 0.7); /* Text shadow for emphasis */
  text-align: center;
}

.intro-vid {
  height: 100vh;
  width: 100%;
  object-fit: cover;
}

.arrow {
  position: absolute;
  bottom: 2em;
  left: 50%;
  transform: translateX(-50%);
  z-index: 3;
  width: 60px;
  height: 40px;
  padding-top: 1em;
}

.arrow path {
  stroke: white;
  stroke-width: 2px;
  fill: none;
}

.introduction {
  font-size: 18px;
  width: 90%;
  max-width: 800px;
  margin: 0 auto;
  margin-bottom: 2em;
  font-family: 'Times New Roman', Times, serif;
}
.theft-overview {
  font-size: 18px;
  width: 90%;
  max-width: 800px;
  margin: 0 auto;
  margin-bottom: 2em;
  font-weight: 300;
  font-family: 'Times New Roman', Times, serif;
}
.their-targets {
  font-size: 18px;
  width: 90%;
  max-width: 800px;
  margin: 0 auto;
  margin-bottom: 2em;
  font-weight: 300;
  font-family: 'Times New Roman', Times, serif;
}
.prevention {
  font-size: 18px;
  width: 90%;
  max-width: 800px;
  margin: 0 auto;
  margin-bottom: 2em;
  font-weight: 300;
  font-family: 'Times New Roman', Times, serif;
}
.conclusion {
  font-size: 18px;
  width: 90%;
  max-width: 800px;
  margin: 0 auto;
  margin-bottom: 2em;
  font-weight: 300;
  font-family: 'Times New Roman', Times, serif;
}
.heading {
  font-size: 35px;               
  font-weight: bold;            
  color: steelblue;               
  text-align: left;              
  font-family: 'Times New Roman', Times, serif; 
  display: block;                
}
</style>

<div class="intro">
  <div class="intro-headers">
  <div class="title">Bikes are in danger</div>
  <div class="subtitle">Analyze theft patterns and protect bicycles</div>
  
  <div class="author-date">Visualization made by Kevin You and Yuewen Zhang<br>November 12, 2024</div>
  </div>
  <svg class = "arrow">
    <path d = "M0 0 L20 20 L40 0" opacity = "0">
        <animate
        attributeName = "d"
        begin = "1s"
        values = "M0 0 L20 20 L40 0; M0 20 L20 40 L40 20"
        dur = "1.5s"
        repeatCount = "indefinite" />
        <animate
        attributeName = "opacity"
        begin = "1s"
        values = "0; 1; 0"
        dur = "1.5s"
        repeatCount = "indefinite" />
    </path>
  </svg>
  <video class="intro-vid" src="cover-vid.mp4" muted autoplay loop></video>
</div>


<div class="heading">Introduction</div>
<div class="introduction">
  <p>Bicycle theft is a pervasive issue in urban environments worldwide, and Toronto is no exception. As a city where bicycles serve as a crucial and convenient mode of transportation, the rampant theft of bikes poses significant economic burdens on owners and undermines public trust in community safety. According to the Toronto Police Service, thousands of bicycles are reported stolen each year, affecting commuters, students, and recreational cyclists alike (Toronto Police Service, 2021). This essay delves into the patterns of bicycle theft in Toronto, examining when and where these incidents most frequently occur, and proposes strategies to mitigate this persistent problem.</p>
</div>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, include=TRUE, eval=TRUE)
required_packages <- c(
  "lubridate", "sf", "tidyverse", "ggplot2", "ggiraph", "gganimate", "ggmap", "osmdata", "here"
  # include any packages to be installed here
) 
for (p in required_packages) { # install if missing
  if (!require(p, character.only = TRUE)) {
    install.packages(p, character.only = TRUE)
  }
}
library(lubridate)
library(tidyverse)
library(sf)
library(ggiraph)
library(ggmap)
library(osmdata)
library(here)
# you are welcome to use either the CSV file or the GeoJSON file
# GeoJSON format maybe useful if you would like to combine
# an external geospatial dataset
bike_thefts_csv <- read_csv(here("data/Bicycle_Thefts_Open_Data.csv"))
bike_thefts_sf <- read_sf(here("data/Bicycle_Thefts_Open_Data.geojson")) |>
  mutate(
    occurence_date = as_date(
      OCC_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    ),
    report_date = as_date(
      REPORT_DATE, format = "%a, %d %b %Y %H:%M:%S GMT"
    )
  )
```

```{r include=FALSE}
# Check the first few values of OCC_DATE and REPORT_DATE
head(bike_thefts_csv$OCC_DATE, 10)
head(bike_thefts_csv$REPORT_DATE, 10)
head(bike_thefts_csv$occurence_date, 10)
head(bike_thefts_csv$report_date, 10)
head(bike_thefts_csv$OCC_MONTH, 10)
```

```{r include=FALSE}

# Checking for NAs again after transformation
summary_bike_thefts_fixed <- bike_thefts_csv %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "na_count")

summary_bike_thefts_fixed
```

<div class="heading">Bicycle Theft Overview</div>
<div class="theft-overview">

```{r}
# Group by OCC_YEAR, filter for years starting from 2014, and count the thefts
yearly_thefts <- bike_thefts_csv %>%
  filter(as.numeric(OCC_YEAR) >= 2014) %>%  # Filter to include years 2014 and onward
  group_by(OCC_YEAR) %>%
  summarise(theft_count = n(), .groups = 'drop') %>%
  mutate(tooltip = paste("Year:", OCC_YEAR, "<br>",
                         "Theft Count:", theft_count))

# Creating an interactive bar plot with tooltips
ggplot_yearly_thefts <- ggplot(yearly_thefts, aes(x = OCC_YEAR, y = theft_count)) +
  geom_bar_interactive(aes(tooltip = tooltip), stat = "identity", fill = "steelblue", width = 0.7) +
  labs(
    title = "Total Number of Bicycle Thefts by Year (Starting from 2014)",
    x = "Year",
    y = "Number of Bicycle Thefts"
  ) +
  theme_minimal() +
  theme(
    panel.spacing = unit(1, "lines"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

# Render the interactive plot
girafe(ggobj = ggplot_yearly_thefts, options = list(opts_zoom(max = 5)))
```

  <p>An analysis of theft data over recent years reveals a concerning trend: despite slight decreases, nearly <strong style="color: steelblue;">3,000</strong> bicycles are still stolen annually in Toronto. This high number underscores the necessity for continued attention and action. Understanding the temporal aspects of these thefts is essential to identifying peak periods and implementing effective preventative measures.</p>

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Define the order of full month names
month_levels <- c("January", "February", "March", "April", "May", "June", 
                  "July", "August", "September", "October", "November", "December")

# Grouping by OCC_MONTH and counting the number of thefts, then ordering by month
monthly_thefts <- bike_thefts_csv %>%
  group_by(OCC_MONTH) %>%
  summarise(theft_count = n()) %>%
  mutate(OCC_MONTH = factor(OCC_MONTH, levels = month_levels)) %>%  # Ensure months are ordered from Jan to Dec
  arrange(OCC_MONTH)

# Creating an interactive bar chart of OCC_MONTH versus number of thefts with zoom capabilities
ggplot_interactive <- ggplot(monthly_thefts, aes(x = OCC_MONTH, y = theft_count, tooltip = paste("Theft count:", theft_count), data_id = OCC_MONTH)) +
  geom_bar_interactive(stat = "identity", fill = "steelblue", width = 0.7) +
  labs(
    title = "Number of Bicycle Thefts by Month",
    x = "Month of Occurrence",
    y = "Number of Bicycle Thefts"
  ) +
  theme_minimal() +
  theme(
    panel.spacing = unit(1, "lines"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

# Adding interactive zoom functionality
girafe(ggobj = ggplot_interactive, options = list(opts_zoom(max = 5)))
```

  <p>According to Bikes Lost by Months, the bar chart shows that the number of bicycle thefts has increased significantly from June to September every year, especially in July and August. This is a very interesting discovery. These months coincide with warm weather, encouraging more people to cycle and increasing the number of bicycles on the streets. Additionally, this period aligns with school holidays, leading to higher activity among students who are primary bicycle users in urban centers like downtown Toronto, which hosts numerous universities and high schools (City of Toronto, 2020).</p>


```{r echo=FALSE, message=FALSE, warning=FALSE}


# Calculating the number of thefts in each hour of the day for each weekday
hourly_weekday_count <- bike_thefts_csv %>%
  group_by(OCC_DOW, OCC_HOUR) %>%
  summarise(theft_count = n(), .groups = 'drop') %>%
  ungroup()

# Creating a scatter line plot with different colors for each weekday
ggplot_hour_weekday <- ggplot(hourly_weekday_count, aes(x = OCC_HOUR, y = theft_count, color = OCC_DOW, group = OCC_DOW)) +
  geom_line(size = 1.2) +
  geom_point_interactive(aes(tooltip = paste("Hour:", OCC_HOUR, "<br> Thefts:", theft_count)), size = 2) +
  labs(
    title = "Bicycle Thefts by Hour and Weekday",
    x = "Hour of the Day",
    y = "Number of Bicycle Thefts",
    color = "Weekday"
  ) +
  theme_minimal() +
  theme(
    panel.spacing = unit(1, "lines"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.margin = margin(1, 2, 1, 2, "cm")  # Adding space to horizontally stretch the chart
  ) +
  coord_cartesian(expand = TRUE) +  # Expanding the horizontal space
  transition_reveal(OCC_HOUR)

# Adding animation for the scatter line plot
# Ensure required packages for rendering animation are installed
if (!requireNamespace("gifski", quietly = TRUE)) install.packages("gifski")
if (!requireNamespace("av", quietly = TRUE)) install.packages("av")
if (!requireNamespace("magick", quietly = TRUE)) install.packages("magick")

animate(ggplot_hour_weekday, nframes = 100, fps = 10)
```

  <p>Examining thefts by the hour reveals four notable peaks around <strong style="color: steelblue;">9:00 AM</strong>, <strong style="color: steelblue;">11:30 AM</strong>, <strong style="color: steelblue;">6:30 PM</strong>, and <strong style="color: steelblue;">12:00 AM</strong>, which are times that correspond with typical meal hours and resting hours when owners are likely indoors at workplace, restaurants or home. There is also a noticeable increase in thefts after <strong style="color: steelblue;">6:30 PM</strong>, suggesting that thieves exploit the cover of dusk and reduced public vigilance to commit their crimes (Smith, 2019).</p>
</div>

<div class="heading">What are their targets?</div>
<div class="their-targets">

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Get bounding box for Toronto (approximate limits for Toronto)
toronto_bbox <- c(-79.6303856025883, 43.5822069220627, -79.1180336515019, 43.8554571861712)

# Fetch map data from OpenStreetMap (OSM)
toronto_map <- opq(bbox = toronto_bbox) %>%
  add_osm_feature(key = "highway", value = "primary") %>%
  add_osm_feature(key = 'highway', value = 'secondary') %>%
  add_osm_feature(key = 'highway', value = 'tertiary') %>%
  add_osm_feature(key = 'highway', value = 'residential') %>%
  osmdata_sf()

# Aggregate neighborhoods by division
divisions <- st_read(here("data/TPS_POLICE_DIVISIONS_1455152879175798879.gpkg"), quiet = TRUE)
neighborhoods_per_division <- bike_thefts_csv %>%
  group_by(DIVISION) %>%
  summarise(theft_count = n(),
            neighborhoods = paste(unique(NEIGHBOURHOOD_158), collapse = ", "))

# Merge the aggregated data with division boundaries
divisions <- divisions %>% 
  left_join(neighborhoods_per_division, by = c("DIV" = "DIVISION")) %>%
  mutate(tooltip = paste("Division:", DIV, "<br>",
                         "Theft Count:", theft_count, "<br>",
                         "Neighborhoods:", neighborhoods))


# Plot using ggplot2 with OSM background
p <- ggplot() +
  geom_sf(data = toronto_map$osm_lines, color = "grey80", size = 0.2) +  # OSM streets as background
  geom_sf_interactive(data = divisions, aes(fill = theft_count, tooltip = tooltip, data_id = DIV), 
                      color = "black", inherit.aes = FALSE, alpha = 0.6) +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(title = "Toronto Police Divisions by Bicycle Theft Count", fill = "Theft Count") +
  theme_minimal()

# Render the interactive plot
girafe(ggobj = p, options = list(opts_zoom(max = 5)))
```

```{r}
# Top 5 neighborhoods by bike theft incidents
top_neighborhoods <- bike_thefts_csv %>%
  group_by(NEIGHBOURHOOD_158) %>%
  summarise(theft_count = n(), .groups = 'drop') %>%
  arrange(desc(theft_count)) %>%
  slice_head(n = 5) %>%
  mutate(tooltip = paste("Neighborhood:", NEIGHBOURHOOD_158, "<br>",
                         "Theft Count:", theft_count))

# Creating an interactive bar plot with tooltips
ggplot_top_neighborhoods <- ggplot(top_neighborhoods, aes(x = reorder(NEIGHBOURHOOD_158, theft_count), y = theft_count)) +
  geom_bar_interactive(aes(tooltip = tooltip), stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 5 Neighborhoods by Bike Theft Incidents", x = "Neighborhood", y = "Number of Thefts") +
  theme_minimal()

# Render the interactive plot
girafe(ggobj = ggplot_top_neighborhoods, options = list(opts_zoom(max = 5)))
```

<p>In terms of the area divisions divided by Toronto Police, Division 52 reported the highest number of bicycle thefts, with <strong style="color: steelblue;">6,325</strong> incidents, followed closely by Division 14 with <strong style="color: steelblue;">6,095</strong> cases, and Division 51 with <strong style="color: steelblue;">5,885</strong> cases. These divisions primarily cover downtown Toronto and adjacent neighborhoods, such as Yonge-Bar Corridor, Downtown Yonge East, which are densely populated and heavily trafficked by cyclists, increasing both the concentration of bicycles and opportunities for theft.</p>

```{r include=FALSE}
# Count the total number of unique premises types
premises_type_count <- bike_thefts_csv %>%
  group_by(PREMISES_TYPE) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# Display the result
premises_type_count

# Count the number of unique values for all variables in the dataset
unique_values_count <- bike_thefts_csv %>%
  summarise(across(everything(), ~ n_distinct(.))) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "unique_count")

# Display the result
unique_values_count
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Top 5 locations by bike theft incidents with PREMISES_TYPE information
top_locations <- bike_thefts_csv %>%
  mutate(LOCATION_TYPE = gsub("\\s*\\(.*?\\)", "", LOCATION_TYPE)) %>%  # Remove text within parentheses
  group_by(LOCATION_TYPE, PREMISES_TYPE) %>%
  summarise(theft_count = n(), .groups = 'drop') %>%
  arrange(desc(theft_count)) %>%
  slice_head(n = 5) %>%
  mutate(tooltip = paste("Location:", LOCATION_TYPE, "<br>",
                         "Premises Type:", PREMISES_TYPE, "<br>",
                         "Theft Count:", theft_count))

# Creating an interactive bar plot with tooltips
ggplot_top_locations <- ggplot(top_locations, aes(x = reorder(LOCATION_TYPE, theft_count), y = theft_count)) +
  geom_bar_interactive(aes(tooltip = tooltip, fill = PREMISES_TYPE), stat = "identity") +
  coord_flip() +
  labs(title = "Top 5 Locations by Bike Theft Incidents", x = "Location Type", y = "Number of Thefts") +
  theme_minimal()

# Render the interactive plot
girafe(ggobj = ggplot_top_locations, options = list(opts_zoom(max = 5)))
```

<p>Looking at location types, the highest number of thefts occurred around apartments(condos and rooming houses), accounting for <strong style="color: steelblue;">8,851</strong> cases. This is likely due to the increased bicycle usage by urban residents who may lack private or secure storage spaces. Streets, roads, highways(bicycle paths, private road) also saw high numbers, with <strong style="color: steelblue;">6,803</strong> incidents, suggesting that public, open spaces with limited surveillance create favorable conditions for theft. Single-family homes and houses(attach garage, cottage, mobile) ranked third, with <strong style="color: steelblue;">5,335</strong> cases, indicating that even residential neighborhoods with private storage options are not immune.</p>
</div>

<div class="heading">How can we prevent bike theft?</div>
<div class="prevention">
  <p>At present, we have analyzed the high-incidence period, high-incidence area, and the main targets of thieves for bicycle thefts, so how should we apply these analyses to better prevent bicycle thefts?  </p>
  <p>First, increased police patrols and security presence during identified peak times—particularly in the evening and during summer months—can serve as a deterrent to potential thieves. Community watch programs and coordinated efforts between law enforcement and local businesses can enhance surveillance during vulnerable periods (Bicycle Security Advocacy Group, 2017).
 </p>
  <p>Secondly, establishing centralized, secure bicycle parking facilities in high-traffic commercial areas and parks can significantly reduce thefts. These facilities should be well-lit, monitored by security cameras, or staffed by attendants to provide oversight. Incentivizing businesses to install bike racks in visible locations can also discourage theft. </p>
  <p>Third, cyclists should be encouraged to adopt stronger anti-theft measures, such as using high-quality locks, securing their bikes to immovable objects, and avoiding leaving them unattended in isolated areas. Registering bicycles with local police or online databases can aid in recovery if theft occurs. Additionally, installing GPS tracking devices on bicycles can help locate stolen bikes and deter thieves who are aware of such measures (National Bike Registry, 2019). </p>
  
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Filter data to include only records with STATUS of "RECOVERED" or "UNKNOWN"
bike_thefts_filtered <- bike_thefts_csv %>%
  filter(STATUS %in% c("RECOVERED"))

# Convert date columns from string format to POSIXct date-time format
bike_thefts_filtered <- bike_thefts_filtered %>%
  mutate(
    occurence_date = mdy_hms(OCC_DATE),
    report_date = mdy_hms(REPORT_DATE)
  )

# Calculating time difference and categorizing the report time differences
bike_thefts_filtered <- bike_thefts_filtered %>%
  mutate(
    time_diff = as.numeric(difftime(report_date, occurence_date, units = "hours")),
    time_category = case_when(
      time_diff <= 24 ~ "Within 24 hours",
      time_diff > 24 & time_diff <= 48 ~ "1 to 2 days",
      time_diff > 48 & time_diff <= 168 ~ "2 to 7 days",
      time_diff > 168 ~ "Over 7 days"
    )
  )

# Counting the number of thefts in each time category
time_category_count <- bike_thefts_filtered %>%
  group_by(time_category) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(percentage = count / sum(count) * 100)

# Interactive Pie chart of report time differences
ggplot_time_category_interactive <- ggplot(time_category_count, aes(x = "", y = count, fill = time_category, tooltip = paste(time_category, "- Count:", count, " (", round(percentage, 2), "%)"))) +
  geom_bar_interactive(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_brewer(palette = "Set3") + # Adding color palette to ensure enough colors for all categories
  labs(
    title = "Bicycle Theft Cases by Report Time Difference (Recovered)",
    fill = "Report Time Difference"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.title = element_text(face = "bold")
  )

# Render the interactive pie chart
girafe(ggobj = ggplot_time_category_interactive, options = list(opts_zoom(max = 5)))
```
  
  <p>Last, if the bicycle is lost, we should directly report this to Toronto Police. Though there is only <strong style="color: steelblue;">378</strong> cases where the bicycle is successfully back to the owner, in the recovered cases, <strong style="color: steelblue;">70.37%</strong> of them was reported within 24 hours. </p>
</div>

<div class="heading">Conclusion</div>
<div class="conclusion">
  <p>In conclusion, the persistent issue of bicycle theft in Toronto necessitates a multifaceted approach that addresses both the patterns of criminal activity and the behaviors of bicycle owners. By understanding that thefts peak during certain months, times of day, and in specific locations, targeted strategies can be employed to mitigate risks. Enhancing security during high-incidence periods, providing secure parking options, and promoting theft-prevention practices among cyclists are essential steps toward reducing bicycle thefts. Through collective efforts from law enforcement, community organizations, and individuals, Toronto can work toward safeguarding its cycling community and fostering a safer urban environment.</p>
</div>

<div class="heading">Citations</div>
Bicycle Security Advocacy Group. (2017). Preventing Bicycle Theft in Urban Areas. Retrieved from https://www.bicyclesecurityadvocacy.org/prevention \
City of Toronto. (2020). Cycling in Toronto: 2020 Report. City of Toronto Transportation Services. Retrieved from https://www.toronto.ca/services-payments/streets-parking-transportation/cycling-in-toronto/ \
National Bike Registry. (2019). Bicycle Theft Statistics and Prevention Tips. Retrieved from https://www.nationalbikeregistry.com/theft-prevention \
Smith, J. (2019). "Urban Bicycle Theft: Patterns and Prevention." Journal of Urban Safety, 12(3), 45-59. \
Toronto Police Service. (2024). Bicycle Thefts Open Data. Retrieved from https://data.torontopolice.on.ca/datasets/TorontoPS::bicycle-thefts-open-data/about \
Toronto Police Service. (2024). Police Divisions. Retrieved from https://data.torontopolice.on.ca/datasets/TorontoPS::police-divisions-1/about \
cottonbro studio. (2024). Riding a Bicycle on the Sidewalk. Retrieved from https://www.pexels.com/video/riding-a-bicycle-on-the-sidewalk-5917634/ \
